---
title: "Aggregating Earth Surface data using GEE and R"
author: "Francesco Pirotti [&lt;francesco.pirotti@unipd.it&gt;](mailto:francesco.pirotti@unipd.it)"
date: "\\#Summer Webinar Series — Sept. 29, 2020"
output: html_document
  # revealjs::revealjs_presentation:
  #   theme: night
  #   #transition: none
  #   self_contained: false
  #   css: slides.css 
  # beamer_presentation:
  #   toc: false
  #   incremental: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Agenda

- From a regular lattice of points create squares to calculate values from aggregating data (e.g. height above sea level of Earth surface)


## Starters

To add functionalities you must load the specific libraries - REMEMBER to make sure that they are installed in your system. If  not you can install in RStudio on **menu => Tools => Install Packages**  
or directly with command  
```{r   eval=F   }
install.packages("NAME OF PACKAGE")
```



```{r  }
library(sf)
library(mapview)
library(raster)
```

### My custom CRS projection Lambert Conical Conformal NOT secant but tangent at lat=45.827 and lon=11.625

```{r}
myproj <- "+proj=lcc +lat_1=45.827  +lat_2=45.827  +lat_0=45.827 +lon_0=11.625 +x_0=4000000 +y_0=2800000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
```

### My points (in lat long) over the regular grid/lattice these points are 666.67 m apart

```{r}
grid.points <- st_read( "data/webinar1_2020_07_29/points.shp" )
```

###  Convert my points from lat long to my custom CRS projection  

```{r}
grid.points.myproj <-   grid.points %>% st_transform(myproj)
```

Run this line to view   

```{r}
  mapview( grid.points.myproj )
```

Trick to create a rhomboid with a certain distance from the point ("radius")   
 half the distance between points (333.33 m)  
```{r}
nodes.buffered<-st_buffer(grid.points.myproj, 333.33, nQuadSegs = 1)
```
 
More complex function the includes another function    

```{r}

st_bbox_by_feature = function(geom) {
  
  ## make sure that object "geom" becomes a geometry object;
  ## geom is your data set with all the single geometries/polygons
  geom2 = st_geometry(geom)
  
  #' Function to:
  #' (a) take a single geometry, 
  #' (b) create a bounding box with st_bbox and 
  #' (c) convert the bbox object to a new geometry (a square)
  f <- function(single.geom) { 
    st_as_sfc( st_bbox(single.geom,), crs=myproj)
  }
  
  #' This line calls the function above LOOPING over each single geometry (lapply)
  #'  in the geom2 object 
  do.call("c", lapply(geom2, f))
}
```

This line calls the above function    
```{r} 
tiles <- st_bbox_by_feature( nodes.buffered )
```

Assign the CRS to the new dataset    
```{r}
tiles <- tiles %>% st_set_crs(myproj)
```

Convert to Latitude and Longitude for ingesting into Google Earth Engine       
```{r}
tiles.latlng <- tiles %>% st_transform("+init=epsg:4326")
```


Save it to a shapefile ... this shapefile will be uploaded to your    
Google Earth Engine space  
```{r}
#st_write(tiles.latlng, "data/tiles.shp" )
```

 Import tiles.shp in GOOGLE EARTH ENGINE AND    
 DO THINGS 
 code shared https://earthengine.googlesource.com/users/2020_Kanan/summer_webinar_series 
 in script
 https://earthengine.googlesource.com/users/2020_Kanan/summer_webinar_series/+/636ab64efd7b09a0d45d3d10a992b8a44bd119c7/webinar1
 AND EXPORT THE RESULTS TO A SHAPEFILE CALLED   
  tiles_withData.shp     
  

Read the file exported from GEE   
```{r}
grid.points.gee <- st_read( "data/webinar1_2020_07_29/tiles_withData.shp" )
```


Create a new attribute column with the interquartile range    
(The difference between p75 and p25)    
```{r}
grid.points.gee$iqr<-grid.points.gee$p75 - grid.points.gee$p25
```

Draw some results!!!!  

Figure 2 - 50th Percentile (median)  

```{r}
pal = mapviewPalette("mapviewSpectralColors")
mapview( grid.points.gee, col.regions = pal(100), zcol = "p50", 
         color=NULL, alpha.regions=0.8  ) 
```

Figure 2 - 10th Percentile  
```{r}
mapview( grid.points.gee, col.regions = pal(100), zcol = "p10", 
         color=NULL, alpha.regions=0.8  ) 
```

Figure 3 - Inter Quartile Range  
```{r}
mapview( grid.points.gee, col.regions = pal(100), zcol = "iqr", 
         color=NULL, alpha.regions=0.8)
```

   
