---
title: "Time-series of RADAR backscatter using GEE and R"
author: "Francesco Pirotti [&lt;francesco.pirotti@unipd.it&gt;](mailto:francesco.pirotti@unipd.it)"
date: "\\#Summer Webinar Series — August xx, 2020"
output:
  html_document:
    code_folding: show
    highlight: tango
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning = F )

```


# Code & Copyleft

```{r, child=c('copyright.Rmd')}
``` 

## Agenda

- Part 2 from webinar related to how Google Earth Engine (GEE) can be used to extract time series per each polygon of areas damaged by VAIA 
- Import extracted data in R CRAN
- Plot time series data and applying a Z-score filter over past N values to detect changes

## Before we begin

- For this Webinar we use  RStudio - if you want to follow along, make sure you have have installed R + RStudio in your PC   
- Make sure you have downloaded the data [download here](data/webinar3/GEEdataFromWebinar2.zip).

## Data
Data come from processing from previous webinar.  
The data is in GeoJSON format, so processing it takes a bit of extra steps as it is different from typical tabular data.
JSON data is unstructured data and very popular in the "BigDATA" era. For further reading:

- 

## R libraries

```{r message=FALSE }
library(sf)
library(mapview)
library(raster)
library(RJSONIO)
```

## Read data   
Data is zipped, but we can extract with R and then read the file.
The first line extracts the file **webinar2vaia.geojson** from the archived dataset in **GEEdataFromWebinar2.zip**...
The second line reads the GeoJSON data.
The last line shows 

```{r}
mydata<-unzip("data/webinar3/GEEdataFromWebinar2.zip")
data.gee <-    fromJSON( mydata )  
head(data.gee)
```

  