---
title: "Webinar 2"
author: "Francesco Pirotti [&lt;francesco.pirotti@unipd.it&gt;](mailto:francesco.pirotti@unipd.it)"
date: "\\#Summer Webinar Series — August 5, 2020"
output:
  html_document:
    code_folding: show
    highlight: tango
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning = F )

```
# Time-series of RADAR backscatter using GEE and R - part 1


## Code & Copyleft

```{r, child=c('copyright.Rmd')}
``` 

## Agenda

The hypothesis is that a healthy/standing forest and a damaged forest will have significantly different backscatter values of RADAR. We will use Sentinel 1 data. For more reading on how RADAR works and Sentinel 1:

- 

 

- Load to Google Earth Engine (GEE) vegetation areas which have been subject to changes: (disturbance by fire, windthrow) 
  - In the past 5 years (Sentinel 1 data are  available since 2014) 
  - I will use areas strongly damaged by VAIA downloadable from the windthrow database   [[@Forzieri2020]](https://essd.copernicus.org/articles/12/257/2020/) linking to an <a href="https://figshare.com/articles/A_spatially-explicit_database_of_wind_disturbances_in_European_forests_over_the_period_2000-2018/9555008" target="_blank">open database</a>  - but you can use your own scenario  
- Extract time series of Sentinel 1 backscatter data - i.e. a statistical aggregate of VV+VH backscatter values inside the polygons
- Import extracted data in R CRAN ()
- Plot time series data and applying a Z-score filter over past N values to detect changes

## Google Earth Engine

- The commented GEE script is included below and is available [here](https://code.earthengine.google.com/?scriptPath=users%2F2020_Kanan%2Fsummer_webinar_series%3Awebinar2).



## Output
The final output from processing in GEE is in GeoJSON format.  Why in GeoJSON and not in Shapefile? 
Because GeoJSON allows non-structured attributes to be linked to each feature (polygon). Let's make an example: our polygons represent an area damaged by VAIA, so it comes with a table, which is  *structured* data, ie. each feature has a row in a table and one or more attributes which are table fields/columns. BUT if we have more complex data, imagine that we want to assign to each polygon the values of 
 
## Conclusion

The next seminar [here](webinar3.html) will be on plotting and processing our extracted data.



  